---
title: "An Analysis of Reported Heights and Weights of Current NFL Players"
author: "Albert Ding"
date: "December 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, echo = FALSE, warning=FALSE, results='hide',message=FALSE}
#Loading libraries:
pacman::p_load(
tidyverse,
data.table,
XML,
RCurl,
httr,
rvest,
magrittr,
benford.analysis,
gridExtra,
knitr,
png,
jpeg
)
```

```{r, echo = FALSE, warning=FALSE, results='hide',message=FALSE}
#Scrape and clean data on NFL rosters

#vector of URLs below:

nfl_teams <- c("arizona-cardinals", "atlanta-falcons", "baltimore-ravens", "buffalo-bills", "carolina-panthers", "chicago-bears", "cincinnati-bengals", "cleveland-browns", "dallas-cowboys", "denver-broncos","detroit-lions","green-bay-packers","houston-texans","indianapolis-colts","jacksonville-jaguars","kansas-city-chiefs","los-angeles-chargers","los-angeles-rams","miami-dolphins", "minnesota-vikings","new-england-patriots","new-orleans-saints","new-york-giants","new-york-jets","oakland-raiders","philadelphia-eagles","pittsburgh-steelers","san-francisco-49ers","seattle-seahawks","tampa-bay-buccaneers","tennessee-titans","washington-redskins") 

result  <- list() #formatting result as list

#for loop to extract tables of rosters below
       
for(i in 1:length(nfl_teams)) {
   url <- paste0("https://www.lineups.com/nfl/roster/", nfl_teams[i])
   tble <- read_html(url) %>%
      html_node("table") %>%
      html_table(fill = TRUE)
   
   result[[i]] <- tble #list of 32 tables - 1 for each team
   
   result[[i]]['Team_name'] = nfl_teams[i] #assigning variable to identify what team each player is on
}

all_teams <- do.call("rbind", result) %>% #merging all 32 tables
    mutate(
    Name1 = sub("^(\\S*\\s+\\S+).*", "\\1", Name)) 
    #creating new column of names because existing has dulplicate of name from picture of player     on website

all_teams$Name <- NULL #deleting original name column

colnames(all_teams)[16] <- "Name" #changing name1 to name

clean1<-all_teams[c(16,1:15)] %>% #reorder columns to put name first
    filter(Height != 0 & Weight != 0) %>% #used to extract inches
    mutate(Height_length = nchar(Height)) %>% #see whether extracting 1 or 2 digits for inches
    mutate(Height_ft = as.numeric(substr(Height, start = 1, stop = 1)),
           Height_in = 
              ifelse(Height_length == 4,
              as.numeric(substr(Height, start = 3, stop = 3)),
              as.numeric(substr(Height, start = 3, stop = 4)))) %>% #extracting inches
    mutate(Height_inches = Height_ft*12 + Height_in) #converting to inches

clean1 <- subset(clean1,Name != "Andrew East") #Removing miscategorized backup long snapper

#Renaming columns for dplyr formatting
colnames(clean1)[colnames(clean1)=="Exp."] <- "Exp"
colnames(clean1)[colnames(clean1)=="Draft Round"] <- "Draft_round"
colnames(clean1)[colnames(clean1)=="Draft Pick"] <- "Draft_pick"

#Selecting relevant columns for final data frame
Final <- select(clean1, Name, Pos, Team_name, Height_inches, Weight, Age, Exp, Rating, Depth, Drafted, Draft_round, Draft_pick, College)

```

## Summary

- <font size="4">Project examines heights and weights of current NFL players for inaccuracies arising from self-reported nature of figures</font>

- <font size="4">Neither player heights, nor player weights follow Benford distribution, but they should both be approximately normal if their disributions are similar to those of the overall population of adult males</font>

- <font size="4">Population further segmented into groups by player position and examined for normality</font>

- <font size="4">For offensive linemen weight distribution, I reproduced the Benford getSuspects function by running  Chi-squared tests on different quantiles to see which ones deviated the most from the expected normal densities</font>

- <font size="4">After significant analytical investigation, I visualize and explore other relationships in the dataset</font>

## Data Scraped from Sports Analytics Website Lineups.com

```{r, echo = FALSE, warning=FALSE, results='hide',message=FALSE}
x <- "https://www.lineups.com/articles/wp-content/uploads/2017/10/nfl-risers-fallers-2.jpg"
download.file(x,'x.jpg', mode = 'wb')
jj <- readJPEG("x.jpg",native=TRUE)
plot(0:1,0:1,type="n",ann=FALSE,axes=FALSE)
rasterImage(jj,0,0,1,1)

```

## Player Heights Don't Follow Benford Distribution and Range from 66 to 81 Inches  

```{r, echo = FALSE, warning=FALSE, results='hide',message=FALSE}
#benford analysis on height 
bfd.height <- benford(Final$Height_inches)

plot(bfd.height)
```

## Player Weights Don't Follow Benford Either and Range from 149 to 362 lbs  

```{r, echo = FALSE, warning=FALSE, results='hide',message=FALSE}
#benford analysis on weight 
bfd.weight <- benford(Final$Weight)

plot(bfd.weight)

```

## Weights Appear Multi-Modal While Heights Have Bell Shape   

```{r, echo = FALSE, warning=FALSE, results='hide',message=FALSE}
Weight_hist1<- ggplot(Final) +
    geom_histogram(aes(x = Weight, y = ..count..),
          binwidth = 5, fill = "grey", color = "black") +
    ggtitle("Distribution of NFL Players Weights\n(5lb Bins)") +
    ylab("\nCount") +
    xlab("Weight in lbs\n") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    coord_flip() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))+ 
    geom_vline(xintercept = mean(Final$Weight), color="red") +
    geom_vline(xintercept = mean(Final$Weight) + sd(Final$Weight), color="blue") +
    geom_vline(xintercept = mean(Final$Weight) - sd(Final$Weight), color="blue")  

Height_hist1<- ggplot(Final) +
    geom_histogram(aes(x = Height_inches, y = ..count..),
          binwidth = 1, fill = "grey", color = "black") +
    ggtitle("Distribution of NFL Players Heights\n(1 inch Bins)") +
    ylab("\nCount") +
    xlab("Height in Inches\n") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    coord_flip() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    geom_vline(xintercept = mean(Final$Height_inches), color="red")+ 
    geom_vline(xintercept = mean(Final$Height_inches) + sd(Final$Height_inches), color="blue") +
    geom_vline(xintercept = mean(Final$Height_inches) - sd(Final$Height_inches), color="blue") 

grid.arrange(Weight_hist1,Height_hist1,nrow=1)
```

## Q-Q Plots Indicate Neither Are Drawn from Normally Distributed Population  

```{r, echo = FALSE, warning=FALSE, message=FALSE}
set.seed(2345)
Final_obs<-nrow(Final)
Final$Height_inches_uniform<- Final$Height_inches+runif(Final_obs, min = -0.5, max = 0.5)

par(mfrow=c(1,2))
qqnorm(Final$Height_inches_uniform, main="Normal Q-Q Plot of Heights\n(Adjusted - See Report)", ylab = "Sample Quantiles - Height in Inches"); qqline(Final$Height_inches_uniform)
qqnorm(Final$Weight,  main="Normal Q-Q Plot of Weights", ylab = "Sample Quantiles - Weight in lbs"); qqline(Final$Weight)

```

## Player Heights Were Adjusted With Randomly Drawn Decimal Endings To Address Discreteness 

```{r, echo = FALSE, warning=FALSE, message=FALSE}
par(mfrow=c(1,2))
qqnorm(Final$Height_inches,  main="Normal Q-Q Plot of Heights\n(Unadjusted)", ylab = "Sample Quantiles - Height in Inches"); qqline(Final$Height_inches)
qqnorm(Final$Height_inches_uniform, main="Normal Q-Q Plot of Heights\n(Adjusted)", ylab = "Sample Quantiles - Height in Inches"); qqline(Final$Height_inches_uniform)

```

## Shapiro-Wilk Test of Normality Agrees They're Not Normally Distributed

- <font size="4">Shapiro-Wilk tests null hypothesis that a sample came from a normally distributed population. p-value is less than 0.05 so we can reject the null hypothesis at 95% LOC that height and weight were drawn from normally distributed population (See report for details)</font>

```{r, echo = FALSE, warning=FALSE, message=FALSE}
par(mfrow=c(1,2))
shapiro.test(Final$Weight)
shapiro.test(Final$Height_inches)

```

## Conclusions from Basic Visualizations of Height and Weights

- <font size="4">Player weights were not normally distributed and heights appear off as well</font>

- <font size="4">Height distribution has elements of discreteness caused by rounding to the nearest inch in reported height values</font>

- <font size="4">There were only 16 different values for height: 66 inches to 81 inches inclusive</font>

- <font size="4">I addressed height discreteness by adding decimal endings from uniform distribution centered at zero and plus / minus half an inch to mimic rounding in real life</font>

- <font size="4">Next, I'll analyze the distribution of heights and weights by position</font>

## Quarterback Height and Weight Distributions Look Bell Shaped

```{r, echo = FALSE, warning=FALSE, message=FALSE}
QB_Final <- filter(Final, Pos == "QB")

Weight_hist_QB<- ggplot(QB_Final) +
    geom_histogram(aes(x = Weight, y = ..count..),
          binwidth = 5, fill = "grey", color = "black") +
    ggtitle("QB Weights\n(5lb Bins)") +
    ylab("\nCount") +
    xlab("Weights\n") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    coord_flip() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    geom_vline(xintercept = mean(QB_Final$Weight), color="red") +
    geom_vline(xintercept = mean(QB_Final$Weight) + sd(QB_Final$Weight), color="blue") +
    geom_vline(xintercept = mean(QB_Final$Weight) - sd(QB_Final$Weight), color="blue")  

Height_hist_QB<- ggplot(QB_Final) +
    geom_histogram(aes(x = Height_inches, y = ..count..),
          binwidth = 1, fill = "grey", color = "black") +
    ggtitle("QB Heights\n(1 inch Bins)") +
    ylab("\nCount") +
    xlab("Heights\n") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    coord_flip() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    geom_vline(xintercept = mean(QB_Final$Height_inches), color="red") +
    geom_vline(xintercept = mean(QB_Final$Height_inches) + sd(QB_Final$Height_inches), color="blue") +
    geom_vline(xintercept = mean(QB_Final$Height_inches) - sd(QB_Final$Height_inches), color="blue")  

grid.arrange(Weight_hist_QB,Height_hist_QB, nrow=1)
```

## QQ Plots for Quarterback Suggest Normal Distribution  

```{r, echo = FALSE, warning=FALSE, message=FALSE}
QB_observations <- nrow(QB_Final)
set.seed(111)
Uniform_decimals_QB <- runif(QB_observations, min =  -0.5, max = 0.5)
QB_Final$Height_inches_uniform<- QB_Final$Height_inches + Uniform_decimals_QB

par(mfrow=c(1,2))
qqnorm(QB_Final$Height_inches_uniform,  main="Normal Q-Q Plot of QB Heights\n(Adjusted)", ylab = "Sample Quantiles - Height in Inches"); qqline(QB_Final$Height_inches_uniform)
qqnorm(QB_Final$Weight,  main="Normal Q-Q Plot of QB Weights", ylab = "Sample Quantiles - Weight in lbs"); qqline(QB_Final$Weight)


```

## Shapiro-Wilk Tests Agrees That Quarterback Distributions May Be Normal

- <font size="4">Both have p-values far greater than 0.05. Notably, if the heights are not adjusted, the p-value becomes small enough to reject the null hypothesis.</font>

```{r, echo = FALSE, warning=FALSE, message=FALSE}
par(mfrow=c(1,2))
shapiro.test(QB_Final$Weight)
shapiro.test(QB_Final$Height_inches_uniform)
```

## In Full Report and Shiny App, We Also Examine Distributions, QQ-Plots, and Tests of Normality for Other Positions

- <font size="4">Omitted from report for sake of compactness and brevity</font>

- <font size="4">We'll look at offensive linemen next for which we apply further testing on to retrieve suspects</font>

## Offensive Line: Distributions Look Bell Shaped But Skewed For Weight 

```{r, echo = FALSE, warning=FALSE, message=FALSE}
OL_Final <- filter(Final, Pos == "OT"|Pos == "OG"|Pos == "OL"|Pos == "C")

Weight_hist_OL<- ggplot(OL_Final) +
    geom_histogram(aes(x = Weight, y = ..count..),
          binwidth = 5, fill = "grey", color = "black") +
    ggtitle("Offensive Lineman Weights\n(5lb Bins)") +
    ylab("\nDensity") +
    xlab("Weights\n") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    coord_flip() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    geom_vline(xintercept = mean(OL_Final$Weight), color="red") +
    geom_vline(xintercept = mean(OL_Final$Weight) + sd(OL_Final$Weight), color="blue") +
    geom_vline(xintercept = mean(OL_Final$Weight) - sd(OL_Final$Weight), color="blue") 

Height_hist_OL<- ggplot(OL_Final) +
    geom_histogram(aes(x = Height_inches, y = ..count..),
          binwidth = 1, fill = "grey", color = "black") +
    ggtitle("Offensive Lineman Heights\n(1 inch Bins)") +
    ylab("\nDensity") +
    xlab("Heights\n") +
    theme(plot.title = element_text(hjust = 0.5)) + 
    coord_flip() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
    geom_vline(xintercept = mean(OL_Final$Height_inches), color="red") +
    geom_vline(xintercept = mean(OL_Final$Height_inches) + sd(OL_Final$Height_inches), color="blue") +
    geom_vline(xintercept = mean(OL_Final$Height_inches) - sd(OL_Final$Height_inches), color="blue")  

grid.arrange(Weight_hist_OL,Height_hist_OL, nrow=1)
```

## Q-Q Plots Indicate Height Close to Normal But Weight is Not

```{r, echo = FALSE, warning=FALSE, message=FALSE}
OL_obs <- nrow(OL_Final)

set.seed(1222)

Uniform_decimals_OL <- runif(OL_obs, min =  -0.5, max = 0.5)

OL_Final$Height_inches_uniform<- OL_Final$Height_inches + Uniform_decimals_OL

par(mfrow=c(1,2))
qqnorm(OL_Final$Weight,  main="Normal Q-Q Plot of O-Line Weight", ylab = "Sample Quantiles - Weight in Pounds"); qqline(OL_Final$Weight)
qqnorm(OL_Final$Height_inches_uniform, main="Normal Q-Q Plot of O-Line Height\n(Adjusted)", ylab = "Sample Quantiles - Height in Inches"); qqline(OL_Final$Height_inches_uniform)
```

## Shapiro-Wilk Tests Agrees Offensive Line Height May be Normal But Not Weight

```{r, echo = FALSE, warning=FALSE, message=FALSE}
shapiro.test(OL_Final$Weight)
shapiro.test(OL_Final$Height_inches_uniform)
```

## Using The Chi-Squared Test Statistic, We'll Find Weight Bins With The Largest Deviations  

- <font size="4">In histogram, weights bucketed in five pound increments with 85 different values in that range so 17 buckets total</font>

- <font size="4">Using each bucket, we can calculate expected area under the curve from pnorm function and multiply by total sample size to get how many samples expected in that quantile</font>

- <font size="4">From there, we can calculate observed minus expected and generate chi-squared values for each bin</font>

- <font size="4">We then select the bins with the largest chi squared test statistic values - this is how the benford analysis package selects suspects</font>


```{r, echo = FALSE, warning=FALSE, results='hide',message=FALSE}
Mean_Weight_OL<-mean(OL_Final$Weight)
SD_Weight_OL<-sd(OL_Final$Weight)

#BUCKET ONE (277-281 lbs)
Bucket_one_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 277 &  OL_Final$Weight <= 281.5,]))
Bucket_one_OL_weight_theoretical<-pnorm(q = 281, mean = Mean_Weight_OL, sd = SD_Weight_OL)*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_one_deviation<- Bucket_one_OL_weight_actual - Bucket_one_OL_weight_theoretical #Observed - actual
OL_Chi_squared_1<-Bucket_one_deviation^2 / Bucket_one_OL_weight_theoretical #Calculating chi squared for bucket

#BUCKET TWO (282-286 lbs)
Bucket_two_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 281.5 &  OL_Final$Weight <= 286.5,]))
Bucket_two_OL_weight_theoretical<-(pnorm(q = 286, mean = Mean_Weight_OL, sd = SD_Weight_OL)-pnorm(q = 282, mean = Mean_Weight_OL, sd = SD_Weight_OL))*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_two_deviation<- Bucket_two_OL_weight_actual - Bucket_two_OL_weight_theoretical #Observed - actual
OL_Chi_squared_2<-Bucket_two_deviation^2 / Bucket_two_OL_weight_theoretical #Calculating chi squared for bucket

#BUCKET THREE (287-291 lbs)
Bucket_three_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 286.5 &  OL_Final$Weight <= 291.5,]))
Bucket_three_OL_weight_theoretical<-(pnorm(q = 291, mean = Mean_Weight_OL, sd = SD_Weight_OL)-pnorm(q = 287, mean = Mean_Weight_OL, sd = SD_Weight_OL))*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_three_deviation<- Bucket_three_OL_weight_actual - Bucket_three_OL_weight_theoretical #Observed - actual
OL_Chi_squared_3<-Bucket_three_deviation^2 / Bucket_three_OL_weight_theoretical #Calculating chi squared for bucket

#BUCKET FOUR (292-296 lbs)
Bucket_four_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 291.5 &  OL_Final$Weight <= 296.5,]))
Bucket_four_OL_weight_theoretical<-(pnorm(q = 296, mean = Mean_Weight_OL, sd = SD_Weight_OL)-pnorm(q = 292, mean = Mean_Weight_OL, sd = SD_Weight_OL))*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_four_deviation<- Bucket_four_OL_weight_actual - Bucket_four_OL_weight_theoretical #Observed - actual
OL_Chi_squared_4<-Bucket_four_deviation^2 / Bucket_four_OL_weight_theoretical #Calculating chi squared for bucket

#BUCKET FIVE (297-301 lbs)
Bucket_five_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 296.5 &  OL_Final$Weight <= 301.5,]))
Bucket_five_OL_weight_theoretical<-(pnorm(q = 301, mean = Mean_Weight_OL, sd = SD_Weight_OL)-pnorm(q = 296, mean = Mean_Weight_OL, sd = SD_Weight_OL))*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_five_deviation<- Bucket_five_OL_weight_actual - Bucket_five_OL_weight_theoretical #Observed - actual
OL_Chi_squared_5<-Bucket_five_deviation^2 / Bucket_five_OL_weight_theoretical #Calculating chi squared for bucket

#BUCKET SIX (302-306 lbs)
Bucket_six_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 301.5 &  OL_Final$Weight <= 306.5,]))
Bucket_six_OL_weight_theoretical<-(pnorm(q = 306, mean = Mean_Weight_OL, sd = SD_Weight_OL)-pnorm(q = 302, mean = Mean_Weight_OL, sd = SD_Weight_OL))*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_six_deviation<- Bucket_six_OL_weight_actual - Bucket_six_OL_weight_theoretical #Observed - actual
OL_Chi_squared_6<-Bucket_six_deviation^2 / Bucket_six_OL_weight_theoretical #Calculating chi squared for bucket

#BUCKET SEVEN (307-311 lbs)
Bucket_seven_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 306.5 &  OL_Final$Weight <= 311.5,]))
Bucket_seven_OL_weight_theoretical<-(pnorm(q = 311, mean = Mean_Weight_OL, sd = SD_Weight_OL)-pnorm(q = 307, mean = Mean_Weight_OL, sd = SD_Weight_OL))*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_seven_deviation<- Bucket_seven_OL_weight_actual - Bucket_seven_OL_weight_theoretical #Observed - actual
OL_Chi_squared_7<-Bucket_seven_deviation^2 / Bucket_seven_OL_weight_theoretical #Calculating chi squared for bucket

#BUCKET EIGHT (312-316 lbs)
Bucket_eight_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 311.5 &  OL_Final$Weight <= 316.5,]))
Bucket_eight_OL_weight_theoretical<-(pnorm(q = 316, mean = Mean_Weight_OL, sd = SD_Weight_OL)-pnorm(q = 312, mean = Mean_Weight_OL, sd = SD_Weight_OL))*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_eight_deviation<- Bucket_eight_OL_weight_actual - Bucket_eight_OL_weight_theoretical #Observed - actual
OL_Chi_squared_8<-Bucket_eight_deviation^2 / Bucket_eight_OL_weight_theoretical #Calculating chi squared for bucket

#BUCKET NINE (317-321 lbs)
Bucket_nine_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 316.5 &  OL_Final$Weight <= 321.5,]))
Bucket_nine_OL_weight_theoretical<-(pnorm(q = 321, mean = Mean_Weight_OL, sd = SD_Weight_OL)-pnorm(q = 317, mean = Mean_Weight_OL, sd = SD_Weight_OL))*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_nine_deviation<- Bucket_nine_OL_weight_actual - Bucket_nine_OL_weight_theoretical #Observed - actual
OL_Chi_squared_9<-Bucket_nine_deviation^2 / Bucket_nine_OL_weight_theoretical #Calculating chi squared for bucket

#BUCKET TEN (322-326 lbs)
Bucket_ten_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 321.5 &  OL_Final$Weight <= 326.5,]))
Bucket_ten_OL_weight_theoretical<-(pnorm(q = 326, mean = Mean_Weight_OL, sd = SD_Weight_OL)-pnorm(q = 322, mean = Mean_Weight_OL, sd = SD_Weight_OL))*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_ten_deviation<- Bucket_ten_OL_weight_actual - Bucket_ten_OL_weight_theoretical #Observed - actual
OL_Chi_squared_10<-Bucket_ten_deviation^2 / Bucket_ten_OL_weight_theoretical #Calculating chi squared for bucket

#BUCKET ELEVEN (327-331 lbs)
Bucket_eleven_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 326.5 &  OL_Final$Weight <= 331.5,]))
Bucket_eleven_OL_weight_theoretical<-(pnorm(q = 331, mean = Mean_Weight_OL, sd = SD_Weight_OL)-pnorm(q = 327, mean = Mean_Weight_OL, sd = SD_Weight_OL))*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_eleven_deviation<- Bucket_eleven_OL_weight_actual - Bucket_eleven_OL_weight_theoretical #Observed - actual
OL_Chi_squared_11<-Bucket_eleven_deviation^2 / Bucket_eleven_OL_weight_theoretical #Calculating chi squared for bucket

#BUCKET TWELVE (332-336 lbs)
Bucket_twelve_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 331.5 &  OL_Final$Weight <= 336.5,]))
Bucket_twelve_OL_weight_theoretical<-(pnorm(q = 336, mean = Mean_Weight_OL, sd = SD_Weight_OL)-pnorm(q = 332, mean = Mean_Weight_OL, sd = SD_Weight_OL))*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_twelve_deviation<- Bucket_twelve_OL_weight_actual - Bucket_twelve_OL_weight_theoretical #Observed - actual
OL_Chi_squared_12<-Bucket_twelve_deviation^2 / Bucket_twelve_OL_weight_theoretical #Calculating chi squared for bucket

#BUCKET THIRTEEN (337-341 lbs)
Bucket_thirteen_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 336.5 &  OL_Final$Weight <= 341.5,]))
Bucket_thirteen_OL_weight_theoretical<-(pnorm(q = 341, mean = Mean_Weight_OL, sd = SD_Weight_OL)-pnorm(q = 337, mean = Mean_Weight_OL, sd = SD_Weight_OL))*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_thirteen_deviation<- Bucket_thirteen_OL_weight_actual - Bucket_thirteen_OL_weight_theoretical #Observed - actual
OL_Chi_squared_13<-Bucket_thirteen_deviation^2 / Bucket_thirteen_OL_weight_theoretical #Calculating chi squared for bucket

#BUCKET FOURTEEN (342-346 lbs)
Bucket_fourteen_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 341.5 &  OL_Final$Weight <= 346.5,]))
Bucket_fourteen_OL_weight_theoretical<-(pnorm(q = 346, mean = Mean_Weight_OL, sd = SD_Weight_OL)-pnorm(q = 341, mean = Mean_Weight_OL, sd = SD_Weight_OL))*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_fourteen_deviation<- Bucket_fourteen_OL_weight_actual - Bucket_fourteen_OL_weight_theoretical #Observed - actual
OL_Chi_squared_14<-Bucket_fourteen_deviation^2 / Bucket_fourteen_OL_weight_theoretical #Calculating chi squared for bucket

#BUCKET FIFTEEN (347-351 lbs)
Bucket_fifteen_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 346.5 &  OL_Final$Weight <= 351.5,]))
Bucket_fifteen_OL_weight_theoretical<-(pnorm(q = 351, mean = Mean_Weight_OL, sd = SD_Weight_OL)-pnorm(q = 347, mean = Mean_Weight_OL, sd = SD_Weight_OL))*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_fifteen_deviation<- Bucket_fifteen_OL_weight_actual - Bucket_fifteen_OL_weight_theoretical #Observed - actual
OL_Chi_squared_15<-Bucket_fifteen_deviation^2 / Bucket_fifteen_OL_weight_theoretical #Calculating chi squared for bucket

#BUCKET SIXTEEN (352-356 lbs)
Bucket_sixteen_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 351.5 &  OL_Final$Weight <= 356.5,]))
Bucket_sixteen_OL_weight_theoretical<-(pnorm(q = 356, mean = Mean_Weight_OL, sd = SD_Weight_OL)-pnorm(q = 352, mean = Mean_Weight_OL, sd = SD_Weight_OL))*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_sixteen_deviation<- Bucket_sixteen_OL_weight_actual - Bucket_sixteen_OL_weight_theoretical #Observed - actual
OL_Chi_squared_16<-Bucket_sixteen_deviation^2 / Bucket_sixteen_OL_weight_theoretical #Calculating chi squared for bucket

#BUCKET SEVENTEEN (357-361 lbs)
Bucket_seventeen_OL_weight_actual<-nrow((OL_Final[OL_Final$Weight >= 356.5 &  OL_Final$Weight <= 361.5,]))
Bucket_seventeen_OL_weight_theoretical<-(1-pnorm(q = 357, mean = Mean_Weight_OL, sd = SD_Weight_OL))*nrow(OL_Final) #Number of observations in set times expected CDF within the bucket
Bucket_seventeen_deviation<- Bucket_seventeen_OL_weight_actual - Bucket_seventeen_OL_weight_theoretical #Observed - actual
OL_Chi_squared_17<-Bucket_seventeen_deviation^2 / Bucket_seventeen_OL_weight_theoretical #Calculating chi squared for bucket

```

## Biggest Suspects Are in the Heaviest Quantile - See Below

```{r, echo = FALSE, warning=FALSE, message=FALSE}
Chisq_OL<-rbind(OL_Chi_squared_1,OL_Chi_squared_2,OL_Chi_squared_3,OL_Chi_squared_4,OL_Chi_squared_5,OL_Chi_squared_6,OL_Chi_squared_7,OL_Chi_squared_8,OL_Chi_squared_9,OL_Chi_squared_10,OL_Chi_squared_11,OL_Chi_squared_12,OL_Chi_squared_13,OL_Chi_squared_14,OL_Chi_squared_15,OL_Chi_squared_16,OL_Chi_squared_17)

plot(Chisq_OL,
     main = "Chi Squared Difference by Weight Bins for Offensive Linemen\n",
     xlab="Weight Bins",
     ylab = "Chi-squared Values",
     pch= 21, bg = 73)
```

## These players are Marcus Cannon of the New England Patriots and Zach Banner of the Pittsburgh Steelers

```{r}
y <- "https://cdn.vox-cdn.com/thumbor/Rce2rfWa4LI49mBpOu5F4yhd2Qw=/0x224:3141x1991/1600x900/cdn.vox-cdn.com/uploads/chorus_image/image/52061629/usa_today_9667961.0.jpeg"
download.file(y,'y.jpg', mode = 'wb')
jj <- readJPEG("y.jpg",native=TRUE)

z <- "https://steelersdepot.com/wp-content/uploads/2018/08/Zach-Banner.jpg"
download.file(z,'z.jpg', mode = 'wb')
kk <- readJPEG("z.jpg",native=TRUE)

par(mfrow=c(1,2))
plot(0:1,0:1,type="n",ann=FALSE,axes=FALSE)
rasterImage(jj,0,0,1,1)
plot(0:1,0:1,type="n",ann=FALSE,axes=FALSE)
rasterImage(kk,0,0,1,1)
```

## Both players Struggled With Weight in the Past and Have Weight-Related Incentives in Contracts

- <font size="4">Both players are listed at varying weights depending on the website and time period and have struggled with being too heavy in the past</font>

- <font size="4">Both also have performance incentives in their contract for making a certain weight during training camp</font>

- <font size="4">Our method of getting suspects via the largest Chi-Squared values appears somewhat successful in finding unusual observations in this test case</font>

## Some Other Selected Relationships Visualized (More in Report/Shiny) - 1

```{r, echo = FALSE, warning=FALSE, results='hide',message=FALSE}
ggplot(Final,aes(x = Height_inches, y = Weight, color = Pos)) +
  geom_point(position = "jitter", alpha = 0.3) +
  geom_smooth(method = lm, se = FALSE) +
  ggtitle("NFL Players Height versus Weight") +
    ylab("Weight in lbs\n") +
    xlab("\nHeight in inches")  +
    theme(plot.title = element_text(hjust = 0.5))
```

## Some Other Selected Relationships Visualized (More in Report/Shiny) - 2

```{r, echo = FALSE, warning=FALSE, results='hide',message=FALSE}
ggplot(Final,aes(x = Height_inches, y = Weight, color = Team_name)) +
  geom_point(position = "jitter", alpha = 0.3) +
  geom_smooth(method = lm, se = FALSE) +
  ggtitle("NFL Players Height versus Weight") +
    ylab("Weight in lbs\n") +
    xlab("\nHeight in inches")  +
    theme(plot.title = element_text(hjust = 0.5))
```

## Some Other Selected Relationships Visualized (More in Report/Shiny) - 3

```{r, echo = FALSE, warning=FALSE, results='hide',message=FALSE}
Final_Depth <- filter(Final, Depth <= 4)

ggplot(na.omit(Final_Depth), aes(x = Height_inches, y = factor(Draft_round))) + 
      geom_jitter(aes(color=factor(Depth)),alpha=0.5) +     
      ggtitle("NFL Players Height versus Draft Round") +
      labs(x = "\nHeight in inches", y = "Draft Round\n", color = "Position on Depth Chart") +
      theme(plot.title = element_text(hjust = 0.5))
```

## Some Other Selected Relationships Visualized (More in Report/Shiny) - 4

```{r, echo = FALSE, warning=FALSE, results='hide',message=FALSE}
ggplot(data = na.omit(Final)) + 
  geom_bar(mapping = aes(x = Team_name, fill = factor(Draft_round)),position = "fill") +
  theme(axis.text.x = element_text(angle = 90, vjust = 1, hjust = 1)) + 
  scale_fill_discrete(labels = 
          c('One','Two','Three','Four','Five','Six','Seven','Undrafted')) +     
      ggtitle("Compositions of Draft Rounds by NFL Team Roster\n") +
      labs(x = "\nTeam Name", y = "Proportion\n", fill = "Draft Round") +
      theme(plot.title = element_text(hjust = 0.5))

```
